---
title: "penguin_analysis"
output: html_document
date: "2024-10-16"
When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

You can also embed plots, for example: ```{r pressure, echo=FALSE}
plot(pressure)

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.


#start of my code
loading in the packages
---
```{r}
library(tidyverse)
library(palmerpenguins)
library(janitor)
library(here)
```



using the here package - this returns the path to your project

```{r}
here::here()
```


penguin data
penguins_raw is a variable
head shows first 6 rows
```{r}
head(penguins_raw)
```


checking column names
- can see there is spaces, brackets and slashes which is not ideal 
- never change the raw data in the excel file, save the raw data and edit it on here
- so, save raw data as csv
```{r}
colnames(penguins_raw)
write.csv(penguins_raw, here("data", "penguins_raw.csv"))
```

remove comments column
- use select and -
```{r}
penguins_raw <- select(penguins_raw, -Comments)
```
-but this is bad code!!!!! we've now overwritten this variable
instead...
- get back penguins raw
- save edits as different names
```{r}
penguins_raw <- read.csv(here("data", "penguins_raw.csv"))
penguins_clean <- select(penguins_raw, -Comments)
```


we can do this for 2 columns at once...
- use select and starts_with to remove delta columns
```{r}
penguins_clean <- select(penguins_raw, -starts_with("delta"))
```
-this is overwriting again!!!

so...
this does everything in one go, we're not overwriting, can just add steps
- this is the best code!!!!! dont look at above
- we are now using piping to make penguins_clean, and then checking its worked using colnames() function
- we added another pipe to clean the names - this makes them all lower case and removes spaces, brackets and slashes
- save it as a csv
```{r}
penguins_raw <- read.csv(here("data", "penguins_raw.csv"))
penguins_clean <- penguins_raw %>%
  select(-Comments) %>% 
  select(-starts_with("delta")) %>%
  clean_names() %>%
  shorten_species()
colnames(penguins_clean)
write.csv(penguins_clean, here("data", "penguins_clean.csv"))
```


we can make our own functions 
- this is basically a quicker way of doing everything we did above!
```{r}
cleaning_penguin_columns <- function(raw_data){
  raw_data %>%
    clean_names() %>%
    select(-comments) %>% 
    select(-starts_with("delta")) %>%
    shorten_species()
}
    
penguins_new <- cleaning_penguin_columns(penguins_raw)
```


can load functions from cleaning.R to our script
- loading all functions in the cleaning.R script
```{r}
source(here("functions", "cleaning.R"))
```


use renv to install packages without using install.packages functions
-snapshot - everuything ive installed is in renv, so anyone else using this code can access all installed packages
```{r}
renv::snapshot()
```

###session 2
```{r}
#flipper_boxplot <- ggplot(
  data = penguins_clean,
  aes(x = species,
      y = flipper_length_mm)) +
    geom_boxplot()

```
subset columns 
```{r}
#penguins_flippers <- select(penguins_clean,c("species", "flipper_length_mm"))
#colnames(penguins_flippers)
```
remove NAs
```{r}
#penguins_flippers <- drop_na(penguins_flippers)
```
but this is bad code - we overwrote penguins_flippers!!
instead, use piping 
```{r}
penguins_flippers <- penguins_clean %>%
  select(species,flipper_length_mm) %>%
  drop_na()
colnames(penguins_flippers)
```
make boxplot again, add colour, add data on top with geom_jitter
then make data same colour as box plot
add another line about position of jitter
then change column names
change colour of plots to make them colour blind friendly - add map to top of code for species colours
```{r}
species_colours <- c("Adelie" = "darkorange",
                     "Chinstrap" = "purple",
                     "Gentoo" = "cyan4")
flipper_boxplot <- ggplot(
  data = penguins_flippers,
  aes(x = species,
      y = flipper_length_mm)) +
    geom_boxplot(aes(color = species),
                 width = 0.3,
                 show.legend = FALSE) +
  geom_jitter(aes(color = species),
              alpha = 0.3,
              show.legend = FALSE,
              position = position_jitter(
                width = 0.2,
                seed = 0)) +
  scale_color_manual(values = species_colours) +
  labs(x = "Species", 
       y = "Flipper Length (mm)") +
  theme_minimal()
flipper_boxplot
```

making function - *unfinished - she said she'd give us the code
```{r}
plot_boxplot <- function(data,
                         x_column,
                         y_column,
                         x_label,
                         y_label, 
                         colour_mapping){
  #drop the na
  data <- data %>%
    drop_na({{y_column}})
  
  #make plot
  ggplot(
    data = data,
    aes(x = {{x_column}},
        y = {{y_column}},
        color = {{}})
  )
```







